Solar wind predictions for the Parker Solar Probe orbit

### 2017-11-05 paperMVVB/sections/content.tex
removed bold text

\section{Introduction}
%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1959 \citep{Gringauz1960,Neugebauer1966}.
%in situ spacecraft
The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in case of Voyager~1 as far away as \SI{140}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}} (\textbf{October} 2017), having even left the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
\textbf{Various} spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with WIND \citep{Lepping1995,Ogilvie1995}\textbf{, SOHO \citep{Domingo1995}}, ACE \citep{Stone1998} and DSCOVR \citep{Burt2012} still orbiting around the L1 point \SI{1.5}{million\,\km} ahead of Earth in the sunward direction. Additional measurements at other solar distances were provided by planetary missions to Venus and Mercury, such as PVO \citep{Colin1980} or MESSENGER \citep{Belcher1991}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind even at polar latitudes \citep{McComas1998}. \textbf{The in situ solar wind measurements closest to the Sun to date were made by} the Helios missions. \textbf{Helios~1, launched in 1974, reached distances of \SI{0.31}{\au}. Helios~2, launched two years later, approached the Sun as close as} \SI{0.29}{\au} \citep{Rosenbauer1977}.
%Solar Probe Plus mission
The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rs), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). In comparison, the ESA Solar Orbiter mission with a planned launch in February 2019 will have its closest perihelia at \SI{0.28}{\au} \citep{Muller2013}.

%PSP objectives and instruments
The key PSP science objectives are to “trace the flow of energy that heats and accelerates the solar corona and solar wind, determine the structure and dynamics of the plasma and magnetic fields at the sources of the solar wind, and explore mechanisms that accelerate and transport energetic particles” as stated in \citet{Fox2015}. To achieve these goals, PSP has four scientific instruments on board: FIELDS for the measurements of magnetic fields and AC/DC electric fields \citep{Bale2016}, SWEAP for the measurements of flux of electrons, protons and alphas \citep{Kasper2016}, IS\sun{}IS for the measurement of solar energetic particles \textbf{(SEPs)} \citep{McComas2016} and WISPR for the measurement of coronal and inner heliospheric structures \citep{Vourlidas2016}.

%CGAUSS and WISPR
The study presented in this paper is undertaken in the Coronagraphic German And US \textbf{SolarProbePlus} Survey (CGAUSS) project, which is the German contribution to the PSP mission as part of the Wide-field Imager for Solar PRobe (WISPR). WISPR will contribute to the PSP science goals by deriving the 3D structure of the solar corona through which the in situ measurements are made to determine the sources of the solar wind. It will provide density power spectra over a wide range of structures (e.g., streamers, pseudostreamers and equatorial coronal holes) for determining the roles of turbulence, waves and pressure-balanced structures in the solar wind. It will also measure the physical properties, such as speed and density jumps of SEP-producing shocks and their \textbf{coronal mass ejection (CME)} drivers as they evolve in the corona and inner heliosphere \citep{Vourlidas2016}.

@@

%definition of sw environment
%section{Solar wind}
Generally two types of solar wind are observed in the heliosphere -- slow and fast streams \citep{Neugebauer1966,Schwenn1983}. Slow solar wind has typical speeds \SI{<400}{\km\per\s} and fast solar wind has speeds \SI{>600}{\km\per\s} \citep[p.~144]{Schwenn1990}. Their different compositions and characteristics indicate different sources and generation processes \citep{McGregor2011a}. Fast streams are found to originate from coronal holes as confirmed by Ulysses' out-of-ecliptic measurements \citep{McComas1998}. The source of slow wind and its eventually different types \citep{Schwenn1983}, is still a subject of controversial discussions because several scenarios are possible to explain its origin from closed magnetic structures in the solar corona, such as intermittent reconnection at the top of helmet streamers and from coronal hole boundaries \citep{Kilpua2016}. The occurrence frequency of these slow and fast streams varies strongly with solar activity and their interactions lead to phenomena such as stream interaction regions and for quasi-stationary coronal source regions to co-rotating interaction regions \citep{Balogh1999}.
Embedded in the slow and fast solar wind streams are transient flows of \textbf{CMEs} -- the faster ones driving shock waves ahead \citep{Gosling1974}. Their rate follows the solar activity cycle and varies in near \SI{1}{\au} measurements between only one CME every couple of days during solar cycle minima up to multiple CMEs observed over several days at times of solar maxima, that is, the CME-associated flow share of the solar wind raises from about \SI{5}{\percent} up to about \SI{50}{\percent} \citep{Richardson2012}.

@@

%sections
Our approach is to obtain analytical representations of the shapes of the solar wind parameter’s frequency distributions in Sect.~\ref{sec:frequency_distribution}, of their solar activity dependence in Sect.~\ref{sec:solar_activity_variations} and of their solar distance scaling in Sect.~\ref{sec:solar_distance_dependency}. The solar wind parameters’ frequency distributions and solar activity dependence is derived from near-Earth solar wind and sunspot number (SSN) time series with a duration of almost five solar cycles. Their distance dependency is derived from Helios solar wind measurements covering more than two third of the distance to the Sun and more than half a solar cycle. From combination of the obtained frequency distributions, SSN dependence functions and solar distance dependence functions a general solar wind model is build in Sect.~\ref{sec:empirical_solar_wind_model}, representing the solar activity and distance behavior. Finally, this empirical model is fed with \textbf{SSN predictions} and extrapolated to PSP's planned orbital positions in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}.

@@

%about OMNI data
The OMNI~2 data set \citep{King2005} combines solar wind magnetic field and plasma data collected by various satellites since 1963, currently by WIND and by ACE. This intercalibrated multi-spacecraft data is time-shifted to the nose of the Earth’s bow shock. The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
%OMNIWeb Data Documentation: http://omniweb.gsfc.nasa.gov/html/ow_data.html
% Acknowledgement to the SPDF OMNIWeb database as the source of data used in publications is requested: "The OMNI data were obtained from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov". Further, for recent years when few sources (IMP 8, Wind, ACE, Geotail) contributed to OMNI, it would be appropriate to also cite the PI's who provided the data to OMNI. Copies of preprints or reprints of OMNI-based publications sent to Natalia Papitashvili (address below) would be appreciated for tracking purposes.
% The best citable reference to OMNI data is J.H. King and N.E. Papitashvili, Solar wind spatial scales in and comparisons of hourly Wind and ACE plasma and magnetic field data, J. Geophys. Res., Vol. 110, No. A2, A02209, 10.1029/2004JA010804.
%
%data time range
In this study the whole hourly data until 31 December 2016 is used, starting from 27 November 1963 (for the temperature from 26 July 1965). The data coverage of the different parameters is in the range \SIrange{67}{74}{\percent},  corresponding to a total duration of 36--40~years.
%OMNI 1963-01-01 data
%B 1963-11-27 14:30
%V 1963-11-27 12:30
%N 1963-11-27 12:30
%T 1965-07-26 00:30
%1963-01-01~00:00--2016-12-31~23:30 all
%19724 d; 54 y\\
% B: 73.91 \%		39.91 y\\
% V: 73.07 \%		39.46 y\\
% N: 70.29 \%		37.96 y\\
% T: 66.73 \%		36.03 y\\
%
%hourly data
\textbf{We note} that a test-comparison of hourly averaged with higher time resolution data for the available shorter time span 1981--2016 did not show significant differences in our results.

%data binning
According to the OMNI data precision and maximal parameter ranges we specify bin sizes of \SI{0.5}{\nT} for the magnetic field strength, \SI{10}{\km\per\s} for the velocity, \SI{1}{\per\cm\cubed} for the density and \SI{10000}{\K} for the temperature. The frequency distributions of the solar wind magnetic field strength, proton velocity, density and temperature are shown in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}.
The solar wind magnetic field strength is in the range \SIrange{0.4}{62}{nT}, the velocity in the range \SIrange{156}{1189}{\km\per\s}, the density in the range \SIrange{0}{117}{\per\cm\cubed}, and the temperature in the range \SIrange{3450}{6.63e6}{\K}, the mean data values are at \SI{6.28}{\nT}, \SI{436}{\km\per\s}, \SI{6.8}{\per\cm\cubed} and \SI{1.05e5}{\K}. These ranges and mean values are as statistically expected from previous analyses of near \SI{1}{\au} solar wind data (e.g., Table~3.3 in \citet[p.~39]{Bothmer2007}).
Much higher or lower peak values at \SI{1}{\au} have been observed in extraordinary events, such as the 23~July 2012 \textbf{CME} with a speed of over \SI{2000}{\km\per\s} and a peak field strength of about \SI{100}{\nT} that was observed by STEREO~A \citep{Russell2013} or the solar wind disappearance event observed in May 1999 with density values even down to \SI{0.2}{\per\cm\cubed} \citep{Lazarus2000}.

@@

	\caption{Resulting fit coefficients from the fitting of the lognormal function (\ref{eq:single_lognormal_fit_function}) to the shape of the solar wind parameters' frequency distributions from near \SI{1}{\au} OMNI hourly data. For the velocity also the fit parameters of the double lognormal function (\ref{eq:double_lognormal_fit_function}) are listed, as well as the median and mean values of the resulting velocity fit. \textbf{The numbers in parentheses are the estimated standard deviations of the fit parameters referred to the corresponding last digits of the quoted value.} \textbf{For each parameter the sum of absolute residuals between data and fit (in percentage of the distribution area) is also listed.}}

@@

		\multicolumn{2}{l}{\textbf{Magnetic field [\si{nT}]}}	&5.661(16)	&6.164(18)	&--	&6.83\\
		\multicolumn{2}{l}{\textbf{Velocity [\SI{e2}{\km\per\s}]}}	&4.085(19)	&4.183(20)	&--	&18.69\\
		\multicolumn{2}{l}{\textbf{Density [\si{\per\cm\cubed}]}}	&5.276(24)	&6.484(34)	&--	&6.48\\
		\multicolumn{2}{l}{\textbf{Temperature [\SI{e4}{\K}]}}	&7.470(17)	&11.301(32)	&--	&5.78\\

@@

		\multirow{2}{*}{\textbf{Velocity}}	&\multicolumn{1}{c}{$W_1$}	&4.89(14)	&5.00(14)	&\multirow{2}{*}{0.504(62)}	&\multicolumn{1}{c}{--}\\
		\multirow{2}{*}{\textbf{[\SI{e2}{\km\per\s}]}~}	&\multicolumn{1}{c}{$W_2$}	&3.68(20)	&3.72(20)	&	&\multicolumn{1}{c}{--}\\

@@

From visual inspection, the resulting fit curves describe the shape of the magnetic field strength, density and temperature distributions well, as can be seen in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}. However, for the velocity the fit function appears not to be as good in describing the measured distribution’s more complex shape around its peak and in the higher velocity range. This also can be inferred from the sum of absolute residuals (SAR) between data and fit, listed in Table~\ref{tab:lognormal_fit_parameters} \textbf{in percentage of the distribution area}, being almost three times larger than those from the other parameters.
\begin{figure*}
	\includegraphics[width=18cm]{figures/histogram_fits_4_a_zoom_paper_pdfplot.pdf}
	\caption{Frequency distributions of the four solar wind parameters and their lognormal fits derived from the hourly OMNI data set. The histograms have bins of \SI{0.5}{\nT}, \SI{10}{\km\per\s}, \SI{1}{\per\cm\cubed} and \SI{10000}{\K}. The fits' median and mean values are indicated as well. The insets show zoomed-in \textbf{views of the high value tails of the distributions. } }
	\label{fig:histogram_fits_4_a_zoom_paper_pdfplot}
\end{figure*}

%double lognormal fitting
%%%%%%%%%%%%%%%%%%%%%%%%%
%justification for compositional lognormal fit
In order to find a better fit result for the velocity distribution, we assume that the velocity distribution can be made up of at least two overlapping branches \citep{McGregor2011b}. Therefore a compositional approach  is chosen by combining two lognormal functions (\ref{eq:single_lognormal_fit_function}), involving more fit variables:
\begin{align}
	W_\text{II}(x) &= c \cdot W_1(x) + (1 -c) \cdot W_2(x)\,.	\label{eq:double_lognormal_fit_function}
\end{align}
The balancing parameter $c$ ensures that the resulting function remains normalized as it represents a probability distribution.
The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution yields the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$) as listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean values of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0	\,.
\end{align}
This more complex fit function is more accurate in describing the velocity's frequency distribution as shown in Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{\textbf{Velocity} frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. \textbf{The inset is a zoomed-in view of the high value tail of the distribution. } }
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
Thus in the following sections we keep the double lognormal ansatz for all velocity frequency fits.

%transition
For the bulk of the solar wind these static lognormal functions describe the parameters' distributions well\textbf{. The abnormally high parameter values in the distribution functions can be attributed to shock/CME events in agreement with the results of the OMNI solar wind investigations by \citet{Richardson2012}.} The simple lognormal fit functions underestimate the \textbf{frequencies in} their high value tails, except for the temperature’s tail which is overestimated, as seen in the insets of Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}\textbf{, this appears to be because CMEs do not come with abnormally high temperatures, but rather with temperatures lower than those of the average solar wind \citep{Forsyth2006}.} The velocity's compositional lognormal fit only slightly overestimates its tail as seen in the inset of Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
The slow and fast part contribute almost equally ($c \approx 0.5$) to the long-term velocity distribution function.

@@

\section{Solar activity dependence of the solar wind frequency distributions}
\label{sec:solar_activity_variations}
In the next step we investigate how the long-term solar wind distribution functions presented in the previous section depend on general solar activity. Therefore we examine their correlation with the \textbf{SSN}, being a commonly used long-term solar activity index, and determine the time lags with the highest correlation coefficients.

%\subsection{SSN correlation}
For the correlations we fit lognormal functions to the frequency distributions as in Sect.~\ref{sec:frequency_distribution}, but implement linear relations to the yearly SSN, allowing shifting of the distribution functions with SSN. For the velocity the approach is different insofar as its two components are kept fixed and instead their balance is modified with the changing SSN. \textbf{Thus we obtain solar activity depending models for the frequency distributions of all four solar wind parameters.}

@@

Yearly medians of the solar wind parameters and the yearly SSN together with the solar cycle number \textbf{are shown in the upper part of Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot}}. The reason for correlating the SSN to the solar wind median values is because the position of a lognormal function is defined by its median. The data are averaged to yearly values to avoid seasonal effects during the Earth’s orbit around the Sun caused by its variations in solar latitude and distance.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{\textbf{Solar} wind parameter yearly \textbf{frequencies (grey shading) with yearly medians (lines)} derived from OMNI data and the yearly SSN from the \citet{sidc} with solar cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}
The solar wind velocity, density and temperature depend on the state of the solar cycle \citep{Schwenn1983}. %density modulation with SSN \citep{Schwenn1983} p.~499\\
For instance the fast solar wind \textbf{occurs at times when polar coronal holes extend to lower latitudes, a typical feature of the declining phase of the solar cycle} as pointed out by \citet[p.~75, Figure~3.52]{Bothmer2007}. Therefore the solar wind velocity, density and temperature maxima exhibit \textbf{time lags relative} to the SSN maxima.

>>>>

\section{Introduction}
%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1959 \citep{Gringauz1960,Neugebauer1966}.
%in situ spacecraft
The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in case of Voyager~1 as far away as \SI{140}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}} in October 2017, having even left the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
Various spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with WIND \citep{Lepping1995,Ogilvie1995}, SOHO \citep{Domingo1995}, ACE \citep{Stone1998} and DSCOVR \citep{Burt2012} still orbiting around the L1 point \SI{1.5}{million\,\km} ahead of Earth in the sunward direction. Additional measurements at other solar distances were provided by planetary missions to Venus and Mercury, such as PVO \citep{Colin1980} or MESSENGER \citep{Belcher1991}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind even at polar latitudes \citep{McComas1998}. The in situ solar wind measurements closest to the Sun to date were made by the Helios missions. Helios~1, launched in 1974, reached distances of \SI{0.31}{\au}. Helios~2, launched two years later, approached the Sun as close as \SI{0.29}{\au} \citep{Rosenbauer1977}.
%Solar Probe Plus mission
The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rs), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). In comparison, the ESA Solar Orbiter mission with a planned launch in February 2019 will have its closest perihelia at \SI{0.28}{\au} \citep{Muller2013}.

%PSP objectives and instruments
The key PSP science objectives are to “trace the flow of energy that heats and accelerates the solar corona and solar wind, determine the structure and dynamics of the plasma and magnetic fields at the sources of the solar wind, and explore mechanisms that accelerate and transport energetic particles” as stated in \citet{Fox2015}. To achieve these goals, PSP has four scientific instruments on board: FIELDS for the measurements of magnetic fields and AC/DC electric fields \citep{Bale2016}, SWEAP for the measurements of flux of electrons, protons and alphas \citep{Kasper2016}, IS\sun{}IS for the measurement of solar energetic particles (SEPs) \citep{McComas2016} and WISPR for the measurement of coronal and inner heliospheric structures \citep{Vourlidas2016}.

%CGAUSS and WISPR
The study presented in this paper is undertaken in the Coronagraphic German And US SolarProbePlus Survey (CGAUSS) project, which is the German contribution to the PSP mission as part of the Wide-field Imager for Solar PRobe (WISPR). WISPR will contribute to the PSP science goals by deriving the 3D structure of the solar corona through which the in situ measurements are made to determine the sources of the solar wind. It will provide density power spectra over a wide range of structures (e.g., streamers, pseudostreamers and equatorial coronal holes) for determining the roles of turbulence, waves and pressure-balanced structures in the solar wind. It will also measure the physical properties, such as speed and density jumps of SEP-producing shocks and their coronal mass ejection (CME) drivers as they evolve in the corona and inner heliosphere \citep{Vourlidas2016}.

@@

%definition of sw environment
%section{Solar wind}
Generally two types of solar wind are observed in the heliosphere -- slow and fast streams \citep{Neugebauer1966,Schwenn1983}. Slow solar wind has typical speeds \SI{<400}{\km\per\s} and fast solar wind has speeds \SI{>600}{\km\per\s} \citep[p.~144]{Schwenn1990}. Their different compositions and characteristics indicate different sources and generation processes \citep{McGregor2011a}. Fast streams are found to originate from coronal holes as confirmed by Ulysses' out-of-ecliptic measurements \citep{McComas1998}. The source of slow wind and its eventually different types \citep{Schwenn1983}, is still a subject of controversial discussions because several scenarios are possible to explain its origin from closed magnetic structures in the solar corona, such as intermittent reconnection at the top of helmet streamers and from coronal hole boundaries \citep{Kilpua2016}. The occurrence frequency of these slow and fast streams varies strongly with solar activity and their interactions lead to phenomena such as stream interaction regions and for quasi-stationary coronal source regions to co-rotating interaction regions \citep{Balogh1999}.
Embedded in the slow and fast solar wind streams are transient flows of CMEs -- the faster ones driving shock waves ahead \citep{Gosling1974}. Their rate follows the solar activity cycle and varies in near \SI{1}{\au} measurements between only one CME every couple of days during solar cycle minima up to multiple CMEs observed over several days at times of solar maxima, that is, the CME-associated flow share of the solar wind raises from about \SI{5}{\percent} up to about \SI{50}{\percent} \citep{Richardson2012}.

@@

%sections
Our approach is to obtain analytical representations of the shapes of the solar wind parameter’s frequency distributions in Sect.~\ref{sec:frequency_distribution}, of their solar activity dependence in Sect.~\ref{sec:solar_activity_variations} and of their solar distance scaling in Sect.~\ref{sec:solar_distance_dependency}. The solar wind parameters’ frequency distributions and solar activity dependence is derived from near-Earth solar wind and sunspot number (SSN) time series with a duration of almost five solar cycles. Their distance dependency is derived from Helios solar wind measurements covering more than two third of the distance to the Sun and more than half a solar cycle. From combination of the obtained frequency distributions, SSN dependence functions and solar distance dependence functions a general solar wind model is build in Sect.~\ref{sec:empirical_solar_wind_model}, representing the solar activity and distance behavior. Finally, this empirical model is fed with SSN predictions and extrapolated to PSP's planned orbital positions in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}.

@@

%about OMNI data
The OMNI~2 data set \citep{King2005} combines solar wind magnetic field and plasma data collected by various satellites since 1963, currently by WIND and by ACE. This intercalibrated multi-spacecraft data is time-shifted to the nose of the Earth’s bow shock. The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
%OMNIWeb Data Documentation: http://omniweb.gsfc.nasa.gov/html/ow_data.html
% Acknowledgement to the SPDF OMNIWeb database as the source of data used in publications is requested: "The OMNI data were obtained from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov". Further, for recent years when few sources (IMP 8, Wind, ACE, Geotail) contributed to OMNI, it would be appropriate to also cite the PI's who provided the data to OMNI. Copies of preprints or reprints of OMNI-based publications sent to Natalia Papitashvili (address below) would be appreciated for tracking purposes.
% The best citable reference to OMNI data is J.H. King and N.E. Papitashvili, Solar wind spatial scales in and comparisons of hourly Wind and ACE plasma and magnetic field data, J. Geophys. Res., Vol. 110, No. A2, A02209, 10.1029/2004JA010804.
%
%data time range
In this study the whole hourly data until 31 December 2016 is used, starting from 27 November 1963 (for the temperature from 26 July 1965). The data coverage of the different parameters is in the range \SIrange{67}{74}{\percent},  corresponding to a total duration of 36--40~years.
%OMNI 1963-01-01 data
%B 1963-11-27 14:30
%V 1963-11-27 12:30
%N 1963-11-27 12:30
%T 1965-07-26 00:30
%1963-01-01~00:00--2016-12-31~23:30 all
%19724 d; 54 y\\
% B: 73.91 \%		39.91 y\\
% V: 73.07 \%		39.46 y\\
% N: 70.29 \%		37.96 y\\
% T: 66.73 \%		36.03 y\\
%
%hourly data
We note that a test-comparison of hourly averaged with higher time resolution data for the available shorter time span 1981--2016 did not show significant differences in our results.

%data binning
According to the OMNI data precision and maximal parameter ranges we specify bin sizes of \SI{0.5}{\nT} for the magnetic field strength, \SI{10}{\km\per\s} for the velocity, \SI{1}{\per\cm\cubed} for the density and \SI{10000}{\K} for the temperature. The frequency distributions of the solar wind magnetic field strength, proton velocity, density and temperature are shown in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}.
The solar wind magnetic field strength is in the range \SIrange{0.4}{62}{nT}, the velocity in the range \SIrange{156}{1189}{\km\per\s}, the density in the range \SIrange{0}{117}{\per\cm\cubed}, and the temperature in the range \SIrange{3450}{6.63e6}{\K}, the mean data values are at \SI{6.28}{\nT}, \SI{436}{\km\per\s}, \SI{6.8}{\per\cm\cubed} and \SI{1.05e5}{\K}. These ranges and mean values are as statistically expected from previous analyses of near \SI{1}{\au} solar wind data (e.g., Table~3.3 in \citet[p.~39]{Bothmer2007}).
Much higher or lower peak values at \SI{1}{\au} have been observed in extraordinary events, such as the 23~July 2012 CME with a speed of over \SI{2000}{\km\per\s} and a peak field strength of about \SI{100}{\nT} that was observed by STEREO~A \citep{Russell2013} or the solar wind disappearance event observed in May 1999 with density values even down to \SI{0.2}{\per\cm\cubed} \citep{Lazarus2000}.

@@

	\caption{Resulting fit coefficients from the fitting of the lognormal function (\ref{eq:single_lognormal_fit_function}) to the shape of the solar wind parameters' frequency distributions from near \SI{1}{\au} OMNI hourly data. For the velocity also the fit parameters of the double lognormal function (\ref{eq:double_lognormal_fit_function}) are listed, as well as the median and mean values of the resulting velocity fit. The numbers in parentheses are the estimated standard deviations of the fit parameters referred to the corresponding last digits of the quoted value. For each parameter the sum of absolute residuals between data and fit (in percentage of the distribution area) is also listed.}

@@

		\multicolumn{2}{l}{Magnetic field [\si{nT}]}	&5.661(16)	&6.164(18)	&--	&6.83\\
		\multicolumn{2}{l}{Velocity [\SI{e2}{\km\per\s}]}	&4.085(19)	&4.183(20)	&--	&18.69\\
		\multicolumn{2}{l}{Density [\si{\per\cm\cubed}]}	&5.276(24)	&6.484(34)	&--	&6.48\\
		\multicolumn{2}{l}{Temperature [\SI{e4}{\K}]}	&7.470(17)	&11.301(32)	&--	&5.78\\

@@

		\multirow{2}{*}{Velocity}	&\multicolumn{1}{c}{$W_1$}	&4.89(14)	&5.00(14)	&\multirow{2}{*}{0.504(62)}	&\multicolumn{1}{c}{--}\\
		\multirow{2}{*}{[\SI{e2}{\km\per\s}]~}	&\multicolumn{1}{c}{$W_2$}	&3.68(20)	&3.72(20)	&	&\multicolumn{1}{c}{--}\\

@@

From visual inspection, the resulting fit curves describe the shape of the magnetic field strength, density and temperature distributions well, as can be seen in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}. However, for the velocity the fit function appears not to be as good in describing the measured distribution’s more complex shape around its peak and in the higher velocity range. This also can be inferred from the sum of absolute residuals (SAR) between data and fit, listed in Table~\ref{tab:lognormal_fit_parameters} in percentage of the distribution area, being almost three times larger than those from the other parameters.
\begin{figure*}
	\includegraphics[width=18cm]{figures/histogram_fits_4_a_zoom_paper_pdfplot.pdf}
	\caption{Frequency distributions of the four solar wind parameters and their lognormal fits derived from the hourly OMNI data set. The histograms have bins of \SI{0.5}{\nT}, \SI{10}{\km\per\s}, \SI{1}{\per\cm\cubed} and \SI{10000}{\K}. The fits' median and mean values are indicated as well. The insets show zoomed-in views of the high value tails of the distributions.}
	\label{fig:histogram_fits_4_a_zoom_paper_pdfplot}
\end{figure*}

%double lognormal fitting
%%%%%%%%%%%%%%%%%%%%%%%%%
%justification for compositional lognormal fit
In order to find a better fit result for the velocity distribution, we assume that the velocity distribution can be made up of at least two overlapping branches \citep{McGregor2011b}. Therefore a compositional approach  is chosen by combining two lognormal functions (\ref{eq:single_lognormal_fit_function}), involving more fit variables:
\begin{align}
	W_\text{II}(x) &= c \cdot W_1(x) + (1 -c) \cdot W_2(x)\,.	\label{eq:double_lognormal_fit_function}
\end{align}
The balancing parameter $c$ ensures that the resulting function remains normalized as it represents a probability distribution.
The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution yields the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$) as listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean values of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0	\,.
\end{align}
This more complex fit function is more accurate in describing the velocity's frequency distribution as shown in Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{Velocity frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset is a zoomed-in view of the high value tail of the distribution.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
Thus in the following sections we keep the double lognormal ansatz for all velocity frequency fits.

%transition
For the bulk of the solar wind these static lognormal functions describe the parameters' distributions well. The abnormally high parameter values in the distribution functions can be attributed to shock/CME events in agreement with the results of the OMNI solar wind investigations by \citet{Richardson2012}. The simple lognormal fit functions underestimate the frequencies in their high value tails, except for the temperature’s tail which is overestimated, as seen in the insets of Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}, this appears to be because CMEs do not come with abnormally high temperatures, but rather with temperatures lower than those of the average solar wind \citep{Forsyth2006}. The velocity's compositional lognormal fit only slightly overestimates its tail as seen in the inset of Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
The slow and fast part contribute almost equally ($c \approx 0.5$) to the long-term velocity distribution function.

@@

\section{Solar activity dependence of the solar wind frequency distributions}
\label{sec:solar_activity_variations}
In the next step we investigate how the long-term solar wind distribution functions presented in the previous section depend on general solar activity. Therefore we examine their correlation with the SSN, being a commonly used long-term solar activity index, and determine the time lags with the highest correlation coefficients.

%\subsection{SSN correlation}
For the correlations we fit lognormal functions to the frequency distributions as in Sect.~\ref{sec:frequency_distribution}, but implement linear relations to the yearly SSN, allowing shifting of the distribution functions with SSN. For the velocity the approach is different insofar as its two components are kept fixed and instead their balance is modified with the changing SSN. Thus we obtain solar activity depending models for the frequency distributions of all four solar wind parameters.

@@

Yearly medians of the solar wind parameters and the yearly SSN together with the solar cycle number are shown in the upper part of Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot}. The reason for correlating the SSN to the solar wind median values is because the position of a lognormal function is defined by its median. The data are averaged to yearly values to avoid seasonal effects during the Earth’s orbit around the Sun caused by its variations in solar latitude and distance.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{Solar wind parameter yearly frequencies (grey shading) with yearly medians (lines) derived from OMNI data and the yearly SSN from the \citet{sidc} with solar cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}
The solar wind velocity, density and temperature depend on the state of the solar cycle \citep{Schwenn1983}. %density modulation with SSN \citep{Schwenn1983} p.~499\\
For instance the fast solar wind occurs at times when polar coronal holes extend to lower latitudes, a typical feature of the declining phase of the solar cycle as pointed out by \citet[p.~75, Figure~3.52]{Bothmer2007}. Therefore the solar wind velocity, density and temperature maxima exhibit time lags relative to the SSN maxima.
